---
title: "Lab Week 8"
author: "Charlene Kormondy"
date: "March 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the packages
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)


```


###Column graph of Texas Oil Spills
```{r}

oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>%  #have to bound it by angled apostrophies bc it has spaces in it
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum (`Net Loss (Barrels)`)) #allows you to apply 1 or more functions to a dataset that is grouped

colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x = Year, y = Loss)) +
  geom_col()

```

###Leaflet plot of spill locations in TX in 2016
```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == "2016") %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326) #not all spatial things will automatically search for longitude latitude, some things need you to do it

leaflet(oil_sf) %>% 
  addTiles %>% 
  addMarkers()

```

##tmap plot with the Texas state shapefile
Sometimes we cant use this base map with leaflet, we need an actual shapefile
```{r}

states <- st_read(dsn = ".", layer = "states") #bc its alrady in our pathway
#now I can treat this information as a data frame. I can use my data frame functions like filter, select, etc. and it will keep spatial geometry stuff with it!

tex_boarder <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326)

plot(tex_boarder)

tm_shape(tex_boarder) + #plot the texas boarder as a polygon
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)

```

###Convert the data to spatial points patters (combination of point data and the bounding window bc we need to have a bounding window to evaluate them within)
we have to convert them into point data with a window that the current functions in r will work with (hopefuly, in the future, you wont have to do this)

```{r}

spill_sp <- as(oil_sf, "Spatial") #converted from simple features back to origional df. but r still doesnt recognize this as data that you want to do point analysis with
spill_ppp <- as(spill_sp, "ppp") #ppp is point pattern p

#state of texas outline is our bounding window:
tx_sp <- as(tex_boarder, "Spatial")
tx_owin <- as(tx_sp, "owin") #now it recognizes it as a window!

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin) #ppp fncn in the spatial stats package for creating point pattern

#4 points were rejected bc they were ouside of the boundary window

```

